Dynamic strategies for yaw and induction control of wind farms based on large-eddy simulation and optimization

### 2018-01-02 reply_to_reviewer2.tex

\documentclass[]{article}

\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage[symbol]{footmisc}
\newcommand{\red}[1]{\textbf{\color{red} #1}}

\newcommand{\revision}[1]{\textbf{#1}}

% Derivatives and differentials
\newcommand{\ds}{~\text{d}\boldsymbol{s}}
\newcommand{\dzeta}{~\text{d}\boldsymbol{\zeta}}
\newcommand{\pder}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\dx}{\text{d}\boldsymbol{x}}
\newcommand{\dt}{\text{d}t}
\newcommand{\ddt}[1]{\frac{\text{d} #1}{\text{d} t}}
% Integrals
\newcommand{\stint}{\int_{0}^{T} \int_{\Omega}}
\newcommand{\sint}{\int_{\Omega}}
\newcommand{\Tint}{\int_{0}^{T}}
\newcommand{\tint}{\int_{0}^{t}}
% Filtered variables
\newcommand{\utilde}{\widetilde{\bs{u}}}
\newcommand{\ptilde}{\widetilde{p}}
\newcommand{\ctnhat}[1]{\widehat{C}_{T,#1}'}
\newcommand{\ctihat}{\widehat{C}_{T,i}'}
\newcommand{\cthat}{\widehat{C}_T'}
% Other symbols
\newcommand{\ctmax}{C_{T,\text{max}}'}
\newcommand{\cti}{C_{T,i}'}
\newcommand{\ctn}[1]{C_{T,#1}'}
\newcommand{\R}{\mathscr{R}}
\newcommand{\J}{\mathscr{J}}
\newcommand{\Jtilde}{\tilde{\mathscr{J}}}
\newcommand{\Jgrad}{\nabla \Jtilde}
\newcommand{\Lagr}{\mathscr{L}}
\newcommand{\eperp}{\bs{e}_\perp}
\newcommand{\eperpi}{\bs{e}_{\perp,i}}
\newcommand{\etransi}{\bs{e}_{\parallel,i}}
\newcommand{\ex}{\bs{e}_x}
\newcommand{\ey}{\bs{e}_y}
\newcommand{\ez}{\bs{e}_z}
\newcommand{\eperpn}[1]{\bs{e}_{\perp,#1}}
\newcommand{\vi}{\frac{1}{A_i} \sint \R_i (\bs{s})~\utilde \cdot \eperpi \ds}
% Operators
\newcommand{\innerproduct}[2]{\bigg( #1, #2 \bigg)}
\newcommand{\innerproductsmall}[2]{\big( #1, #2 \big)}
\newcommand{\sumturbines}{\sum_{i=1}^{N_t}}
\newcommand{\diracdelta}{{\delta}}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

\paragraph{16. Reviewer} \textit{Line 596: the "and" is missing in "... yaw angle equation \_\_\_ the adjoint gradient"}

\paragraph{Response} Thanks for pointing this out. Fixed.

\hrulefill

\paragraph{17. Reviewer} \textit{In Eq. A13 $\rho_i$, $\gamma_i$ and $\chi_i$ should be defined. Currently they are defined in A.4.}

\paragraph{Response} Thanks for pointing this out. The manuscript has been updated as follows (\red{line XX, p YY}):

``
adjoint variables \revision{$\bs{q}^* = [\bs{\xi}, \pi, \sigma_1, \dots, \sigma_{N_t}, \eta_1, \dots, \eta_{N_t}, \rho_1, \dots, \rho_{N_t}, \bs{\gamma_1}, \dots, \bs{\gamma}_{N_t}, \chi_1, \dots \chi_{N_t}]$, which serve as Lagrange multipliers. The additional variables $\rho_i, \bs{\gamma}_i$, and $\chi_i$ are the adjoint counterparts of $\R_i, \eperpi$, and $V_i$, resulting from including the definitions of the latter as explicit constraints in (A10 - A12). The resulting Lagrangian is expressed as follows:}
''

### 2018-01-02 MuntersMeyers_DynamicStrategiesControl_revision_WM01.tex
R2C17

{\small
\begin{alignat}{4}
& \underset{\bs{\varphi}, \bs{q}}{\text{minimize}}  & \qquad  \J(\bs{\varphi}, \bs{q}) &= - \Tint \sum_{i=1}^{N_t} \frac{1}{2} C_{P,i}'~V_i^3 A_i \dt  & \\
& \text{s.t.}                      			&         \frac{\partial \utilde}{\partial t} + \big(\utilde \cdot \nabla \big)~ \utilde &= - \nabla (\ptilde + \ptilde_\infty) / \rho - \nabla \cdot \boldsymbol{\tau}_{sgs} + \sum_{i=1}^{N_t} - \frac{1}{2} \ctihat~V_i^2 \R_i(\bs{x})~\eperpi \quad  & \text{in } \Omega \times (0,T], \\
&                                                   &        \nabla \cdot \utilde&=0 									        & \text{in } \Omega \times (0,T],\\
&                                                   &        \tau \ddt{\ctihat}&=\cti - \ctihat 								& i=1\dots N_t~\text{in } (0,T],\\
&                                                   &        \ddt{\theta_i}&=\omega_i											& i=1\dots N_t~\text{in } (0,T],\\
&                                                   &        \R_i(\bs{x})&= \sint G(\bs{s} - \bs{x})~H\big[D/2 - \vert\vert \bs{s} - \bs{t}_i \vert\vert_2\big]~\diracdelta \big[ (\bs{s} - \bs{t}_i)\cdot \eperpi \big] \ds											& i=1\dots N_t~\text{in } (0,T],\\
&                                                   &        \eperpi&=\bs{e}_x \cos \theta_i  + \bs{e}_y \sin \theta_i											& i=1\dots N_t~\text{in } (0,T],\\
&                                                   &        V_i&= \vi											& i=1\dots N_t~\text{in } (0,T].
\end{alignat}
}

\noindent Further, we define the Lagrangian of this optimization problem by adding the state constraints, with shorthand notation $\bs{B}(\bs{\varphi}, \bs{q})$, to the cost functional through inner product with the adjoint variables $\bs{q}^* = [\bs{\xi}, \pi, \sigma_1, \dots, \sigma_{N_t}, \eta_1, \dots, \eta_{N_t}]$, which serve as Lagrange multipliers:

@@

\noindent The adjoint equations in \eqref{eq:adjoint_momentum} -- \eqref{eq:adjoint_eta} can then be found by postulating vanishing gradients of the Lagrangian as  $\mathscr{L}_{\bs{q}_j}(\delta \bs{q}_j) = (\partial \mathscr{L}/\partial \bs{q}_j, \delta \bs{q}_j) = 0$ with respect to each of the state variables $\bs{q}_j$ in $\bs{q} = [\utilde, \ptilde, \ctnhat{1}, \dots, \ctnhat{N_t}, \theta_1, \dots, \theta_{N_t}]$ for all possible variations $\delta \bs{q}_j$ (see, e.g. Refs. \cite{troltzsch,borzinschulz,hinze2008optimization}).  Further, given that these equations are satisfied, the gradient of the reduced cost functional can be found by expressing the sensitivity of the Lagrangian to the control variables as $\Lagr_{\varphi_j}(\delta \varphi_j)$ for the control parameters $\varphi_j$ in $\bs{\varphi} = [\ctn{1}, \dots, \ctn{N_t}, \omega_1, \dots, \omega_{N_t}]$. 
Since the reduced formulation of the optimization problem by definition implies that $B\big(\bs{\varphi}, \bs{q}(\bs{\varphi})\big)=0$, it can be found that the reduced cost functional $\Jtilde$ and the Lagrangian $\Lagr$ are in fact equivalent \cite{hinze2008optimization}, i.e.

>>>>

{\small
\begin{alignat}{4}
& \underset{\bs{\varphi}, \bs{q}}{\text{minimize}}  & \qquad  \J(\bs{\varphi}, \bs{q}) &= - \Tint \sum_{i=1}^{N_t} \frac{1}{2} C_{P,i}'~V_i^3 A_i \dt  & \\
& \text{s.t.}                      			&         \frac{\partial \utilde}{\partial t} + \big(\utilde \cdot \nabla \big)~ \utilde &= - \nabla (\ptilde + \ptilde_\infty) / \rho - \nabla \cdot \boldsymbol{\tau}_{sgs} + \sum_{i=1}^{N_t} - \frac{1}{2} \ctihat~V_i^2 \R_i(\bs{x})~\eperpi \quad  & \text{in } \Omega \times (0,T], \\
&                                                   &        \nabla \cdot \utilde&=0 									        & \text{in } \Omega \times (0,T],\\
&                                                   &        \tau \ddt{\ctihat}&=\cti - \ctihat 								& i=1\dots N_t~\text{in } (0,T],\\
&                                                   &        \ddt{\theta_i}&=\omega_i											& i=1\dots N_t~\text{in } (0,T],\\
&                                                   &        \R_i(\bs{x})&= \sint G(\bs{s} - \bs{x})~H\big[D/2 - \vert\vert \bs{s} - \bs{t}_i \vert\vert_2\big]~\diracdelta \big[ (\bs{s} - \bs{t}_i)\cdot \eperpi \big] \ds											& i=1\dots N_t~\text{in } (0,T], \label{eq:rconstraint}\\
&                                                   &        \eperpi&=\bs{e}_x \cos \theta_i  + \bs{e}_y \sin \theta_i											& i=1\dots N_t~\text{in } (0,T],\\
&                                                   &        V_i&= \vi											& i=1\dots N_t~\text{in } (0,T]. \label{eq:vconstraint}
\end{alignat}
}

\noindent Further, we define the Lagrangian of this optimization problem by adding the state constraints, with shorthand notation $\bs{B}(\bs{\varphi}, \bs{q})$, to the cost functional through inner product with the adjoint variables \revision{$\bs{q}^* = [\bs{\xi}, \pi, \sigma_1, \dots, \sigma_{N_t}, \eta_1, \dots, \eta_{N_t}, \rho_1, \dots, \rho_{N_t}, \bs{\gamma_1}, \dots, \bs{\gamma}_{N_t}, \chi_1, \dots \chi_{N_t}]$, which serve as Lagrange multipliers. The additional variables $\rho_i, \bs{\gamma}_i$, and $\chi_i$ are the adjoint counterparts of $\R_i, \eperpi$, and $V_i$, resulting from including the definitions of the latter as explicit constraints in (\ref{eq:rconstraint} - \ref{eq:vconstraint}). The resulting Lagrangian is expressed as follows:}

@@

\noindent The adjoint equations in \eqref{eq:adjoint_momentum} -- \eqref{eq:adjoint_eta} can then be found by postulating vanishing gradients of the Lagrangian as  $\mathscr{L}_{\bs{q}_j}(\delta \bs{q}_j) = (\partial \mathscr{L}/\partial \bs{q}_j, \delta \bs{q}_j) = 0$ with respect to each of the state variables \revision{$\bs{q}_j$ in $\bs{q} = [\utilde, \ptilde, \ctnhat{1}, \dots, \ctnhat{N_t}, \theta_1, \dots, \theta_{N_t}, \R_1, \dots, \R_{N_t}, \eperpn{1}, \dots, \eperpn{N_t}, V_1, \dots, V_{N_t}]$} for all possible variations $\delta \bs{q}_j$ (see, e.g. Refs. \cite{troltzsch,borzinschulz,hinze2008optimization}).  Further, given that these equations are satisfied, the gradient of the reduced cost functional can be found by expressing the sensitivity of the Lagrangian to the control variables as $\Lagr_{\varphi_j}(\delta \varphi_j)$ for the control parameters $\varphi_j$ in $\bs{\varphi} = [\ctn{1}, \dots, \ctn{N_t}, \omega_1, \dots, \omega_{N_t}]$. 
Since the reduced formulation of the optimization problem by definition implies that $B\big(\bs{\varphi}, \bs{q}(\bs{\varphi})\big)=0$, it can be found that the reduced cost functional $\Jtilde$ and the Lagrangian $\Lagr$ are in fact equivalent \cite{hinze2008optimization}, i.e.
